### 2014-07-24 analysis/figures/snr.tex
Figure layout latex added

\documentclass[]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}
\tableofcontents
\section{SSVEP}
\subsection{SSVEP 10 Hz}
\clearpage
\subsubsection{SSVEP 10 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 1} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_13_41_ssvep_10-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_07_13_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_14_00_01_ssvep_10-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_14_19_09_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB Junction} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_14_40_36_ssvep_10-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_16_00_33_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Cervical Ventebral Dens} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_15_10_16_ssvep_10-crop.pdf} &
\includegraphics[scale=0.18,draft]{../ssvep/matlab_data/_Tue_06_05_2014_11_13_41_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 2} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_15_46_00_ssvep_10-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_16_36_31_ssvep_10-crop.pdf}
\end{tabular}
\caption{Basilar tip, Mid-Basilar, VB Junction, Cervical Ventebral Dens (not for rabbit 10), Basiliar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 10 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_21_50_ssvep_10-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_12_15_ssvep_ctr_10-crop.pdf} \\
\end{tabular}
\caption{Basilar tip. Rabbit 9 basilar tip has the experiment start/stop twice.}
\end{center}
\end{figure}
\subsubsection{SSVEP 10 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1.7cm}Unspecified} & 
\includegraphics[scale=0.30]{../ssvep/matlab_data/_Thu_15_05_2014_17_15_37_ssvep_10-crop.pdf}
\end{tabular}
\caption{Only measured for rabbit 10. Unspecified location.}
\end{center}
\end{figure}
\subsection{SSVEP 40 Hz}
\clearpage
\subsubsection{SSVEP 40 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 1} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_08_22_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_14_02_01_ssvep_40-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_14_20_24_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB Junction} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_14_41_46_ssvep_40-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_16_02_44_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Cervical Ventebral Dens} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_15_11_25_ssvep_40-crop.pdf} &
\includegraphics[scale=0.18,draft]{../ssvep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 2} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_15_48_24_ssvep_50-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_16_38_47_ssvep_40-crop.pdf}
\end{tabular}
\caption{Basilar tip, Mid-Basilar, VB Junction, Cervical Ventebral Dens (not for rabbit 10), Basiliar tip. \_Tue\_06\_05\_2014\_15\_48\_24\_ssvep appears to be mislabeled as a 50 Hz trial.}
\end{center}
\end{figure}
\subsubsection{SSVEP 40 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_23_01_ssvep_40-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40-crop.pdf} \\
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 40 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1.7cm}Unspecified} & 
\includegraphics[scale=0.30]{../ssvep/matlab_data/_Thu_15_05_2014_17_18_01_ssvep_40-crop.pdf}
\end{tabular}
\caption{Only measured for rabbit 10. Unspecified location.}
\end{center}
\end{figure}
\subsection{SSVEP 50 Hz}
\clearpage
\subsubsection{SSVEP 50 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 1} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_16_01_ssvep_50-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_09_32_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_14_03_09_ssvep_50-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_14_21_40_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB Junction} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_14_43_48_ssvep_50-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_16_04_56_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Cervical Ventebral Dens} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_15_12_40_ssvep_50-crop.pdf} &
\includegraphics[scale=0.18,draft]{../ssvep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 2} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_15_49_35_ssvep_40-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_16_40_58_ssvep_50-crop.pdf}
\end{tabular}
\caption{Basilar tip, Mid-Basilar, VB Junction, Cervical Ventebral Dens (not for rabbit 10), Basiliar tip. \_Tue\_06\_05\_2014\_16\_38\_47\_ssvep appears to be mislabeled as a 40 Hz trial.}
\end{center}
\end{figure}
\subsubsection{SSVEP 50 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip} & 
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_24_15_ssvep_50-crop.pdf} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_14_36_ssvep_ctr_50-crop.pdf} \\
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 50 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1.7cm}Unspecified} & 
\includegraphics[scale=0.30]{../ssvep/matlab_data/_Thu_15_05_2014_17_20_11_ssvep_50-crop.pdf}
\end{tabular}
\caption{Only measured for rabbit 10. Unspecified location}
\end{center}
\end{figure}
\section{SSAEP}
\subsection{SSAEP 12 Hz}
\subsubsection{SSAEP 12 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 1} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_31_23_ssaep_12-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_28_28_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_14_08_46_ssaep_12-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_14_24_26_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB Junction} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_14_50_29_ssaep_12-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_16_07_59_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Cervical Ventebral Dens} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_15_18_04_ssaep_12-crop.pdf} &
\includegraphics[scale=0.18,draft]{../ssvep/matlab_data/_Tue_06_05_2014_11_13_41_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 2} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_15_55_07_ssaep_12-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_16_54_15_ssaep_12-crop.pdf}
\end{tabular}
\caption{Basilar tip, Mid-Basilar, VB Junction, Cervical Ventebral Dens (not for rabbit 10), Basiliar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 12 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_39_46_ssaep_12-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_23_40_ssaep_ctr_12-crop.pdf} \\
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 12 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1.7cm}Unspecified} & 
\includegraphics[scale=0.30]{../ssaep/matlab_data/_Thu_15_05_2014_17_08_12_ssaep_12-crop.pdf} &
\end{tabular}
\caption{Only measured for rabbit 10. Unspecified location}
\end{center}
\end{figure}
\subsection{SSAEP 42 Hz}
\subsubsection{SSAEP 42 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 1} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_36_06_ssaep_42-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_29_48_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_14_09_59_ssaep_42-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_14_25_41_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB Junction} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_14_51_59_ssaep_42-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_16_10_09_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Cervical Ventebral Dens} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_15_19_16_ssaep_42-crop.pdf} &
\includegraphics[scale=0.18,draft]{../ssvep/matlab_data/_Tue_06_05_2014_11_13_41_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 2} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_15_56_42_ssaep_42-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_16_56_24_ssaep_42-crop.pdf}
\end{tabular}
\caption{Basilar tip, Mid-Basilar, VB Junction, Cervical Ventebral Dens (not for rabbit 10), Basiliar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 42 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_41_03_ssaep_42-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_24_55_ssaep_ctr_42-crop.pdf} \\
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 42 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1.7cm}Unspecified} & 
\includegraphics[scale=0.30]{../ssaep/matlab_data/_Thu_15_05_2014_17_10_30_ssaep_42-crop.pdf}
\end{tabular}
\caption{Only measured for rabbit 10. Unspecified location}
\end{center}
\end{figure}
\subsection{SSAEP 86 Hz}
\subsubsection{SSAEP 86 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 1} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_37_22_ssaep_86-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_31_02_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_14_11_09_ssaep_86-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_14_26_54_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB Junction} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_14_53_21_ssaep_86-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_16_12_19_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Cervical Ventebral Dens} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_15_20_29_ssaep_86-crop.pdf} &
\includegraphics[scale=0.18,draft]{../ssvep/matlab_data/_Tue_06_05_2014_11_13_41_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip 2} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_15_57_52_ssaep_86-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_16_58_34_ssaep_86-crop.pdf}
\end{tabular}
\caption{Basilar tip, Mid-Basilar, VB Junction, Cervical Ventebral Dens (not for rabbit 10), Basiliar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 86 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 9 & Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Basilar Tip} & 
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_42_15_ssaep_86-crop.pdf} &
\includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86-crop.pdf} \\
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 86 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1.7cm}Unspecified} & 
\includegraphics[scale=0.30]{../ssaep/matlab_data/_Thu_15_05_2014_17_12_38_ssaep_86-crop.pdf}
\end{tabular}
\caption{Only measured for rabbit 10. Unspecified location}
\end{center}
\end{figure}
\end{document}