### 2014-07-24 analysis/figures/snr_tip.tex

\documentclass[]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{float}
\usepackage{graphicx}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}
\tableofcontents
\section{SSVEP}
\subsection{SSVEP 10 Hz}
\clearpage
\subsubsection{SSVEP 10 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_13_41_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_03_34_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_29_43_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_13_17_59_ssvep_10-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_07_13_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_13_20_12_ssvep_10-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 10 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_21_50_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_11_10_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_22_36_ssvep_10-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_13_25_33_ssvep_10-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_12_15_ssvep_ctr_10-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_13_25_05_ssvep_ctr_10-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 10 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}???} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_17_15_37_ssvep_10-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsection{SSVEP 40 Hz}
\subsubsection{SSVEP 40 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_04_50_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_30_51_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_13_19_10_ssvep_40-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_08_22_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_13_21_25_ssvep_40-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 40 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_23_01_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_12_21_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_23_46_ssvep_40-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_13_26_44_ssvep_40-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_13_26_16_ssvep_ctr_40-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 40 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}???} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_17_18_01_ssvep_40-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsection{SSVEP 50 Hz}
\subsubsection{SSVEP 50 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_16_01_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_05_58_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_32_00_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_13_20_19_ssvep_50-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_09_32_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_13_22_37_ssvep_50-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 50 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_11_24_15_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_13_29_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_12_24_55_ssvep_50-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} &
\includegraphics[scale=0.18]{../ssvep/matlab_data/_Tue_06_05_2014_13_27_53_ssvep_50-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_12_14_36_ssvep_ctr_50-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_13_27_27_ssvep_ctr_50-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSVEP 50 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}???} & \includegraphics[scale=0.18]{../ssvep/matlab_data/_Thu_15_05_2014_17_20_11_ssvep_50-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\section{SSAEP}
\subsection{SSAEP 12 Hz}
\subsubsection{SSAEP 12 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_31_23_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_55_37_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl}                              & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_12_38_56_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_13_03_56_ssaep_12-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_28_28_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_13_35_22_ssaep_12-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 12 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_39_46_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_51_43_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl}                              & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_12_43_40_ssaep_12-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_13_09_08_ssaep_12-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_23_40_ssaep_ctr_12-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_13_40_19_ssaep_ctr_12-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 12 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}???} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_17_08_12_ssaep_12-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsection{SSAEP 42 Hz}
\subsubsection{SSAEP 42 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_36_06_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_56_47_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl}                              & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_12_40_15_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_13_06_20_ssaep_42-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_29_48_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_13_36_42_ssaep_42-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 42 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_41_03_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_52_54_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl}                              & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_12_44_55_ssaep_42-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_13_10_18_ssaep_42-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_24_55_ssaep_ctr_42-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_13_41_32_ssaep_ctr_42-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 42 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}???} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_17_10_30_ssaep_42-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsection{SSAEP 86 Hz}
\subsubsection{SSAEP 86 Hz (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_37_22_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_58_00_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl}                              & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_12_41_28_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_13_07_35_ssaep_86-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_31_02_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_13_38_08_ssaep_86-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 86 Hz (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 9 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_42_15_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Hub}                      & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_11_54_01_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Ag/AgCl}                              & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_12_46_06_ssaep_86-crop.pdf} \\
\rotatebox{90}{\hspace{0.2cm}Guidewire @ 10cm from catheter tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Tue_06_05_2014_13_12_42_ssaep_86-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}Guidewire @ Tip} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Coil}            & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_13_42_44_ssaep_ctr_86-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\subsubsection{SSAEP 86 Hz (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
& Rabbit 10 \\
\rotatebox{90}{\hspace{1cm}???} & \includegraphics[scale=0.18]{../ssaep/matlab_data/_Thu_15_05_2014_17_12_38_ssaep_86-crop.pdf}
\end{tabular}
\caption{Basilar tip.}
\end{center}
\end{figure}
\end{document}