### 2014-07-24 analysis/figures/vep.tex

\documentclass[]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}

\tableofcontents

\section{Rabbit 9}
\subsection{Rabbit 9 (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
Rabbit 9 & Early & Late \\
\rotatebox{90}{\hspace{0.0cm}Basilar Tip, wire@tip} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep__late-crop.pdf} \\
%\rotatebox{90}{\hspace{0cm}Basilar Tip, guidewire@hub} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_12_07_06_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_12_07_06_vep__late-crop.pdf} \\
%\rotatebox{90}{\hspace{0cm}Basilar Tip, Ag/AgCl}   & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_12_33_09_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_12_33_09_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{0cm}Basilar Tip, wire@10cm}& \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_13_21_31_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_13_21_31_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Mid-Basilar}                & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_14_04_18_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_14_04_18_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}VB junction}                & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_14_44_57_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_14_44_57_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{0cm}Cervical Vertebral Dens}    & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_15_13_50_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_15_13_50_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{1cm}Basilar Tip}                & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_15_50_43_vep_-crop.pdf} & \includegraphics[scale=0.2]{../vep/matlab_data/_Tue_06_05_2014_15_50_43_vep__late-crop.pdf} 
\end{tabular}
\caption{Rabbit 9 Experiment. TODO: write down data analysis conditions (filter - high, low, notch - parameters)}
\end{center}
\end{figure}

\subsection{Rabbit 9 (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
Rabbit 9 & Early & Late \\
\rotatebox{90}{\hspace{0cm}Basilar Tip, guidewire@tip} & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{0cm}Basilar Tip, guidewire@hub} & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_12_14_37_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_12_14_37_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{0cm}Basilar Tip, Ag/AgCl}       & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_12_26_04_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_12_26_04_vep__late-crop.pdf} \\
\rotatebox{90}{\hspace{0cm}Basilar Tip, guidewire@10cm}& \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_13_30_15_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Tue_06_05_2014_13_30_15_vep__late-crop.pdf} 
\end{tabular}
\caption{Rabbit 9 Live Control. TODO: write down data analysis conditions (filter - high, low, notch - parameters)}
\end{center}
\end{figure}

\section{Rabbit 10}
\subsection{Rabbit 10 (Experiment)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
Rabbit 10 & Early & Late \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip, guidewire} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__late-crop.pdf} \\ % Position: Basilar Tip, Electrode: Guidewire
\rotatebox{90}{\hspace{0.5cm}Basilar Tip, coil} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_13_13_47_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_13_13_47_vep__late-crop.pdf} \\ % Position: Basilar Tip, Electrode: Coil
\rotatebox{90}{\hspace{1cm}Mid-Basilar} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__late-crop.pdf} \\ % Position: Mid-Basilar, Electrode: X-pedion 10 guidewire
\rotatebox{90}{\hspace{1cm}Vertebro-basilar} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep__late-crop.pdf} \\ % Position: Vertebro-basilar (VB) Junction, Electrode: X-pedion 10 guidewire
%\includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_16_32_25_vep_-crop.pdf} \\ % Position: Basilar Tip, Electrode: X-pedion 10 guidewire
%\includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_16_43_27_vep_-crop.pdf} \\ % Position: Basilar Tip, Electrode: X-pedion 10 guidewire
\rotatebox{90}{\hspace{1cm}Basilar Tip} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep__late-crop.pdf} \\ % Position: Basilar Tip, Electrode: X-pedion 10 guidewire
\end{tabular}
\caption{Rabbit 10. TODO: write down data analysis conditions (filter - high, low, notch - parameters)}
\end{center}
\end{figure}

\subsection{Rabbit 10 (Live Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
Rabbit 10 & Early & Late \\
\rotatebox{90}{\hspace{0cm}Basilar Tip, guidewire@tip} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr_late-crop.pdf} \\
\rotatebox{90}{\hspace{0.6cm}Basilar Tip, coil}          & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_13_28_55_vep_ctr-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_13_28_55_vep_ctr_late-crop.pdf} 
\end{tabular}
\caption{Rabbit 10 Live Control. TODO: write down data analysis conditions (filter - high, low, notch - parameters)}
\end{center}
\end{figure}

\subsection{Rabbit 10 (Dead Control)}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
Rabbit 10 & Early & Late \\
\rotatebox{90}{\hspace{1cm}Unknown} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_17_22_22_vep_-crop.pdf} & \includegraphics[scale=0.23]{../vep/matlab_data/_Thu_15_05_2014_17_22_22_vep__late-crop.pdf}
\end{tabular}
\caption{Rabbit 10 Dead Control. TODO: write down data analysis conditions (filter - high, low, notch - parameters)}
\end{center}
\end{figure}
%PSD is computed with overlapping window -- how does this affect results?
%fourier transform vs power spectral density
%options for computing in matlab
%why does overlap produce better curves, peaks where you expect them to be
%
%two interpretations of snr
% - noise is just power
% - noise is variance/std of power
%
%mark off multiples of 60 Hz
%
%\section{PSD vs. DFT}
%\subsection{PSD}
%http://en.wikipedia.org/wiki/Welch's\_method
%http://en.wikipedia.org/wiki/Bartlett\%27s\_method
%http://en.wikipedia.org/wiki/Periodogram
%
%\begin{align*}
%\frac{T}{2}a &= \sum f(t)\cos(kt)
%\frac{T}{2}b &= \sum f(t)\sin(kt)
%r &= \sqrt{a^2 + b^2}
%\end{align*}
%
%\subsection{FFT}
%\begin{align*}
%X_k &= \sum_{n=0}^{N-1}x_ne^{-i2\pi k\frac{n}{N}} = c+di \\
%&= \sum_{n=0}^{N-1}x_n\left(\cos(2\pi k\frac{n}{N}) + i\sin(2\pi k\frac{n}{N})\right)
%r &= |X_k| = \sqrt{c^2+d^2}
%\end{align*}

\end{document}