### 2014-07-27 analysis/figures/vep_all.tex
Latex for all VEP curves

\documentclass[]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\pagestyle{plain}

% scale:
% 0.161
% 0.126
% 0.126

% 0.161
% 0.126
% 0.126
\begin{document}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
%& VEP On Response \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__all.png} \\
\rotatebox{90}{\hspace{0.5cm}Mid-Basilar} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__all.png} \\
\rotatebox{90}{\hspace{0.5cm}Vertebro-basilar} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep__all.png} \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep__all.png}
\end{tabular}
\caption{Rabbit 10 experimental VEP.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{c}
%VEP On Response & VEP Off Response \\
\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr_all.png}
\end{tabular}
\caption{Rabbit 10 live control VEP.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
%& VEP On Response \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep__all.png} \\
\rotatebox{90}{\hspace{0.5cm}Mid-Basilar} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Tue_06_05_2014_14_04_18_vep__all.png} \\
\rotatebox{90}{\hspace{0.5cm}Vertebro-basilar} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Tue_06_05_2014_14_44_57_vep__all.png} \\
\rotatebox{90}{\hspace{0cm}Cervical Vertebral Dens} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Tue_06_05_2014_15_13_50_vep__all.png} \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../vep/matlab_data/_Tue_06_05_2014_15_50_43_vep__all.png}
\end{tabular}
\caption{Rabbit 9 experimental responses.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{c}
%VEP On Response & VEP Off Response \\
\includegraphics[scale=0.161]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep__all.png}
\end{tabular}
\caption{Rabbit 9 live control.}
\end{center}
\end{figure}
\end{document}