### 2014-08-07 analysis/figures/paper_figures.tex
VEP off now shown, and leftmost plot has labels.

\documentclass[]{article}

\usepackage[margin=1.0in]{geometry}
%\usepackage[paperwidth=5.5in, paperheight=8.5in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=1.0]{surgical_methods.pdf}
\caption{Surgical approach, anatomy imaging.} % Figure 1
\end{center}
\end{figure}

\setcounter{figure}{1}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc|cc}
& Rabbit 10 & Rabbit 9 & Rabbit 10 Control & Rabbit 9 Control \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep_-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__off_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep__off-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Thu_15_05_2014_14_26_54_ssaep_86_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Tue_06_05_2014_11_37_22_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Tue_06_05_2014_11_42_15_ssaep_86-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Thu_15_05_2014_14_20_24_ssvep_40_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.10]{../ssavep/matlab_data/_Tue_06_05_2014_11_23_01_ssvep_40-crop.pdf}}
\end{tabular}
\caption{At Basilar Tip for rabbit 9; at mid-basilar for rabbit 10.} % Figure 2
\end{center}
\end{figure}

\setlength{\tabcolsep}{2pt}
\begin{figure}[H]
\begin{center}
\begin{tabular}{ccccc|cc}
& Basilar Tip & Mid-Basilar & VB Junction & Basilar Tip & Live Control & Dead Control \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_17_22_22_vep_-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__off_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../vep/matlab_data/_Thu_15_05_2014_17_22_22_vep__off-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_12_31_02_ssaep_86_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_14_26_54_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_16_12_19_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_16_58_34_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_17_12_38_ssaep_86-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_12_08_22_ssvep_40_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_14_20_24_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_16_02_44_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_16_38_47_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.08]{../ssavep/matlab_data/_Thu_15_05_2014_17_18_01_ssvep_40-crop.pdf}}
\end{tabular}
\caption{Rabbit 10 position dependence.} % Figure 3
\end{center}
\end{figure}
\setlength{\tabcolsep}{6pt}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
\includegraphics[scale=0.5]{impedance.pdf}
\end{tabular}
\caption{Impedence characterization.} % Figure 4
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{ccc}
\raisebox{-0.5\height}{contaminated raw trace + cardiac} &
\raisebox{-0.5\height}{\includegraphics[scale=1.0]{TODO.pdf}} \\
\raisebox{-0.5\height}{VEP aligned contaminated EKG} &
\raisebox{-0.5\height}{\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__all_qrs.png}} \\
\raisebox{-0.5\height}{cardiac aligned avg vep + EKG} &
\raisebox{-0.5\height}{\includegraphics[scale=0.150]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__qrs-crop.pdf}} \\
\raisebox{-0.5\height}{vep aligned cleaned} &
\raisebox{-0.5\height}{\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__all.png}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.161]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep_.pdf}} \\
\raisebox{-0.5\height}{raw cleaned} &
\raisebox{-0.5\height}{\includegraphics[scale=1.0]{TODO.pdf}}
\end{tabular}
\caption{Cardiac contamination and cleaning.} % Figure 5
\end{center}
\end{figure}

\end{document}