### 2014-08-07 analysis/figures/difference.tex

\documentclass[]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}

\begin{figure}[H]
\begin{center}
\raisebox{-0.5\height}{\includegraphics[scale=0.2]{brain.pdf}}
\hspace{0.2cm}
\raisebox{-0.5\height}{\includegraphics[scale=1.0]{../config/legend.pdf}} \\
\vspace{0.5cm}
\begin{tabular}{ccc}
& SSVEP 40 Hz & SSAEP 86 Hz \\
\rotatebox{90}{\hspace{1.0cm}Rabbit 10} &
\includegraphics[scale=0.200]{../ssavep/matlab_data/10rabbit_may_15_2014_ssaep_15_12.pdf} &
\includegraphics[scale=0.200]{../ssavep/matlab_data/10rabbit_may_15_2014_ssvep_11_8.pdf} \\
\rotatebox{90}{\hspace{1.0cm}Rabbit 9} &
\includegraphics[scale=0.200]{../ssavep/matlab_data/9rabbit_may_6_2014_ssvep_8_11.pdf} &
\includegraphics[scale=0.200]{../ssavep/matlab_data/9rabbit_may_6_2014_ssaep_14_17.pdf}
\end{tabular}
\caption{}
\end{center}
\end{figure}

\end{document}