### 2014-08-07 analysis/figures/experimental_spectral_content.tex

\documentclass[]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}

\begin{figure}[H]
\begin{center}
\raisebox{-0.5\height}{\includegraphics[scale=0.2]{brain.pdf}}
\hspace{0.2cm}
\raisebox{-0.5\height}{\includegraphics[scale=1.0]{../config/legend.pdf}} \\
\vspace{0.5cm}
\begin{tabular}{ccc}
& SSVEP 40 Hz & SSAEP 86 Hz \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_12_08_22_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_12_31_02_ssaep_86_experimental.pdf} \\
\rotatebox{90}{\hspace{0.5cm}Mid-Basilar} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_14_20_24_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_14_26_54_ssaep_86_experimental.pdf} \\
\rotatebox{90}{\hspace{0.5cm}Vertebro-basilar} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_16_02_44_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_16_12_19_ssaep_86_experimental.pdf} \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_16_38_47_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_16_58_34_ssaep_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 10 experimental responses - spectral content during experimental.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cccc}
SSVEP 40 Hz & SSAEP 86 Hz \\
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 10 live control - spectral content during experimental.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cccc}
SSVEP 40 Hz & SSAEP 86 Hz \\
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_17_18_01_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_17_12_38_ssaep_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 10 dead control - spectral content during experimental.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cccc}
SSVEP 40 Hz & SSAEP 86 Hz \\
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_09_52_53_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Thu_15_05_2014_10_04_16_ssaep_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 10 baseline - spectral content during experimental.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\raisebox{-0.5\height}{\includegraphics[scale=0.2]{brain.pdf}}
\hspace{0.2cm}
\raisebox{-0.5\height}{\includegraphics[scale=1.0]{../config/legend.pdf}} \\
\vspace{0.5cm}
\begin{tabular}{ccc}
& SSVEP 40 Hz & SSAEP 86 Hz \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_11_37_22_ssaep_86_experimental.pdf} \\
\rotatebox{90}{\hspace{0.5cm}Mid-Basilar} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_14_02_01_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_14_11_09_ssaep_86_experimental.pdf} \\
\rotatebox{90}{\hspace{0.5cm}Vertebro-basilar} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_14_41_46_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_14_53_21_ssaep_86_experimental.pdf} \\
%\rotatebox{90}{\hspace{0cm}Cervical Vertebral Dens} &
%\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_15_11_25_ssvep_40_experimental.pdf} &
%\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_15_20_29_ssaep_86_experimental.pdf} \\
\rotatebox{90}{\hspace{0.5cm}Basilar Tip} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_15_48_24_ssvep_50_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_15_57_52_ssaep_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 9 experimental responses - spectral content during experimental.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
SSVEP 40 Hz & SSAEP 86 Hz \\
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_11_23_01_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_11_42_15_ssaep_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 9 live control - spectral content during experimental.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
SSVEP 40 Hz & SSAEP 86 Hz \\
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_09_52_56_ssvep_40_experimental.pdf} &
\includegraphics[scale=0.161]{../ssavep/matlab_data/_Tue_06_05_2014_10_08_48_ssaep_86_experimental.pdf}
\end{tabular}
\caption{Rabbit 9 baseline - spectral content during experimental.}
\end{center}
\end{figure}

\end{document}