### 2014-08-07 analysis/figures/mid_basilar.tex

\documentclass[]{article}

\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\pagestyle{plain}

\begin{document}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.3]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__aligned_full.png}
\caption{Rabbit 10 mid-basilar aligned VEP data.}
\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.3]{../vep/matlab_data/_Tue_06_05_2014_14_04_18_vep__aligned_full.png}
\caption{Rabbit 9 mid-basilar aligned VEP data.}
\end{center}
\end{figure}

The maximum voltage in the first 0.4 seconds is aligned.

\end{document}