### 2014-08-08 analysis/figures/paper_figures.tex
Documentation of figures

\documentclass[]{article}

\usepackage[margin=0.5in]{geometry}
%\usepackage[paperwidth=5.5in, paperheight=8.5in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{multirow}

\pagestyle{plain}

\begin{document}

%% Rabbit 10 = Subject 1
%% Rabbit 9 = Subject 2
\begin{figure}[H]
\begin{center}
\includegraphics[scale=1.0]{surgical_methods.pdf}
\caption{Surgical approach, anatomy imaging.}
\label{fig:surgical}
% Figure 1
\end{center}
\end{figure}

\setlength{\tabcolsep}{2pt}
\begin{figure}[H]
\begin{center}
\includegraphics[]{../config/legend2.pdf} % TODO: vertical center
\hspace{0.3cm}
\begin{tabular}{c|cc|cc}
& \multicolumn{2}{c|}{Experiment} & \multicolumn{2}{c}{Control} \\
& Subject 1 & Subject 2 & Subject 1 & Subject 2 \\
\hline
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep_-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__off_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Tue_06_05_2014_11_17_10_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Tue_06_05_2014_11_25_22_vep__off-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_14_26_54_ssaep_86_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Tue_06_05_2014_11_37_22_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Tue_06_05_2014_11_42_15_ssaep_86-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_14_20_24_ssvep_40_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Tue_06_05_2014_11_14_51_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Tue_06_05_2014_11_23_01_ssvep_40-crop.pdf}}
\end{tabular}
\caption{Performance of the electrodes at the optimal location for the endovascular location compared with the performance on the live control trials. The SSAEP and SSVEP experimental trials for Subject 1 are shown for the mid-basilar location. All other performances are shown at the basilar tip position. In each of the experiments, all experimental trials elicited a much larger response than the control experiments, which indicates that true non-artifactual responses were recorded. In addition, the endovascular electrode achieved superior signals compared to the scalp electrodes for the VEP and SSAEP experiments.}
\label{fig:both}
% Figure 2
\end{center}
\end{figure}
\setlength{\tabcolsep}{6pt}

\setlength{\tabcolsep}{2pt}
\begin{figure}[H]
\begin{center}
\begin{tabular}{c|cccc|cc}
         & \multicolumn{4}{c|}{Experiment}                       & \multicolumn{2}{c}{Control} \\
         &             &             &             &             & Live         & Dead \\
\hline
Location & Basilar Tip & Mid-Basilar & VB Junction & Basilar Tip & Basilar Tip  & Basilar Tip \\
Time     & 11:57:42    & 14:13:26    & 15:54:54    & 16:47:47    & 12:15:47     & 17:22:22 \\
Time     & 0:00:00 & 2:15:44 & 3:57:12 & 4:50:05 & 0:18:05 & 5:24:40 \\ % TODO: time relative to anesthesia? relative to surgery
\hline
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep_-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_17_22_22_vep_-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_11_57_42_vep__off_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_15_54_54_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_16_47_47_vep__off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_12_15_47_vep_ctr_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_17_22_22_vep__off-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_12_31_02_ssaep_86_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_14_26_54_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_16_12_19_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_16_58_34_ssaep_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_12_26_26_ssaep_ctr_86-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_17_12_38_ssaep_86-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_12_08_22_ssvep_40_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_14_20_24_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_16_02_44_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_16_38_47_ssvep_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_12_13_26_ssvep_ctr_40-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../ssavep/matlab_data/_Thu_15_05_2014_17_18_01_ssvep_40-crop.pdf}}
\end{tabular}
\caption{Demonstration of position dependence for Subject 1. The performance of the endovascular electrode varies as its location is changed. The performance is consistent between the two basilar tip trials. }
\label{fig:position}
% Figure 3
\end{center}
\end{figure}
\setlength{\tabcolsep}{6pt}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc}
\includegraphics[scale=0.5]{impedance.pdf}
\end{tabular}
\caption{Impedance characterization.}
\label{fig:impedance}
% Figure 4
\end{center}
\end{figure}

%\begin{figure}[H]
%\begin{center}
%\begin{tabular}{cc}
%\multirow{2}{*}{
%\begin{tabular}{ccc}
%Raw Trace & VEP Aligned & \\
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__start-crop.pdf}} &
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__cardiac_labelled-crop.pdf}} &
%\\
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__end-crop.pdf}} &
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__labelled-crop.pdf}} &
%\end{tabular}} &
%\raisebox{-0.5\height}{cardiac aligned avg vep + EKG} \\
%& \raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__qrs-crop.pdf}} \\
%%\raisebox{-0.5\height}{contaminated raw trace + cardiac} &
%%\raisebox{-0.5\height}{VEP aligned contaminated EKG} &
%%\raisebox{-0.5\height}{vep aligned cleaned} &
%%\raisebox{-0.5\height}{raw cleaned} &
%%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__end-crop.pdf}}
%\end{tabular}

%\begin{figure}[H]
%\begin{center}
%\begin{tabular}{cccc}
%& Raw Trace & VEP Aligned & cardiac aligned avg vep + EKG \\
%\raisebox{-0.5\height}{\rotatebox{90}{Contaminated}} &
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__start-crop.pdf}} &
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__cardiac_labelled-crop.pdf}} &
%\raisebox{-0.5\height}{\multirow{2}{*}{\includegraphics[scale=0.64]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__qrs-crop.pdf}}} \\
%\raisebox{-0.5\height}{\rotatebox{90}{Cleaned}} &
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__end-crop.pdf}} &
%\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__labelled-crop.pdf}} &
%\end{tabular}

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc|cc}
\raisebox{-0.5\height}{\multirow{3}{*}{
\begin{tabular}{c}
Cardiac-Aligned  Measurements \vspace{0.7cm} \\
\includegraphics[scale=0.64]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__qrs-crop.pdf}
\end{tabular}
}} &
& Raw Trace & VEP-Aligned \\
\hline
&
\raisebox{-0.5\height}{\rotatebox{90}{Contaminated}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__start-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__cardiac_labelled-crop.pdf}} \\
&
\raisebox{-0.5\height}{\rotatebox{90}{Cleaned}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../qrs/matlab_data/_Thu_15_05_2014_14_13_26_vep__end-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../vep/matlab_data/_Thu_15_05_2014_14_13_26_vep__labelled-crop.pdf}} \\
\end{tabular}
\caption{Cardiac contamination and cleaning. 1076 trials for qrs mean}
\label{fig:cardiac}
% Figure 5
\end{center}
\end{figure}

\section*{Figure Information}
This section is information about which trials where used for the figures and some information about the trials.
\begin{table}[H]
\begin{center}
\begin{tabular}{c|cc|cc}
& \multicolumn{2}{c|}{Experiment} & \multicolumn{2}{c}{Control} \\
& Subject 1 & Subject 2 & Subject 1 & Subject 2 \\
\hline
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_11\_57\_42\_vep\_\_labelled-crop.pdf &
../vep/matlab\_data/\_Tue\_06\_05\_2014\_11\_17\_10\_vep\_-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_12\_15\_47\_vep\_ctr-crop.pdf &
../vep/matlab\_data/\_Tue\_06\_05\_2014\_11\_25\_22\_vep\_-crop.pdf \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_11\_57\_42\_vep\_\_off\_labelled-crop.pdf &
../vep/matlab\_data/\_Tue\_06\_05\_2014\_11\_17\_10\_vep\_\_off-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_12\_15\_47\_vep\_ctr\_off-crop.pdf &
../vep/matlab\_data/\_Tue\_06\_05\_2014\_11\_25\_22\_vep\_\_off-crop.pdf \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_14\_26\_54\_ssaep\_86\_labelled-crop.pdf &
../ssavep/matlab\_data/\_Tue\_06\_05\_2014\_11\_37\_22\_ssaep\_86-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_12\_26\_26\_ssaep\_ctr\_86-crop.pdf &
../ssavep/matlab\_data/\_Tue\_06\_05\_2014\_11\_42\_15\_ssaep\_86-crop.pdf \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_14\_20\_24\_ssvep\_40\_labelled-crop.pdf &
../ssavep/matlab\_data/\_Tue\_06\_05\_2014\_11\_14\_51\_ssvep\_40-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_12\_13\_26\_ssvep\_ctr\_40-crop.pdf &
../ssavep/matlab\_data/\_Tue\_06\_05\_2014\_11\_23\_01\_ssvep\_40-crop.pdf
\end{tabular}
\end{center}
\caption{Figure \ref{fig:both} files used to generate plot.}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{c|cccc|cc}
         & \multicolumn{4}{c|}{Experiment}                       & \multicolumn{2}{c}{Control} \\
         &             &             &             &             & Live         & Dead \\
\hline
Location & Basilar Tip & Mid-Basilar & VB Junction & Basilar Tip & Basilar Tip  & Basilar Tip \\
Time     & 11:57:42    & 14:13:26    & 15:54:54    & 16:47:47    & 12:15:47     & 17:22:22 \\
Time     & 0:00:00 & 2:15:44 & 3:57:12 & 4:50:05 & 0:18:05 & 5:24:40 \\ % TODO: time relative to anesthesia? relative to surgery
\hline
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_11\_57\_42\_vep\_\_labelled-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_14\_13\_26\_vep\_-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_15\_54\_54\_vep\_-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_16\_47\_47\_vep\_-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_12\_15\_47\_vep\_ctr-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_17\_22\_22\_vep\_-crop.pdf \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_11\_57\_42\_vep\_\_off\_labelled-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_14\_13\_26\_vep\_\_off-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_15\_54\_54\_vep\_\_off-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_16\_47\_47\_vep\_\_off-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_12\_15\_47\_vep\_ctr\_off-crop.pdf &
../vep/matlab\_data/\_Thu\_15\_05\_2014\_17\_22\_22\_vep\_\_off-crop.pdf \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_12\_31\_02\_ssaep\_86\_labelled-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_14\_26\_54\_ssaep\_86-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_16\_12\_19\_ssaep\_86-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_16\_58\_34\_ssaep\_86-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_12\_26\_26\_ssaep\_ctr\_86-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_17\_12\_38\_ssaep\_86-crop.pdf \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_12\_08\_22\_ssvep\_40\_labelled-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_14\_20\_24\_ssvep\_40-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_16\_02\_44\_ssvep\_40-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_16\_38\_47\_ssvep\_40-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_12\_13\_26\_ssvep\_ctr\_40-crop.pdf &
../ssavep/matlab\_data/\_Thu\_15\_05\_2014\_17\_18\_01\_ssvep\_40-crop.pdf
\end{tabular}
\caption{Figure \ref{fig:position} files used to generate plot.}
\label{fig:position}
% Figure 3
\end{center}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{cc|cc}
Filename & Index & VEP On & VEP Off \\
\hline
\_Tue\_06.05.2014\_11\%3A17\%3A10\_vep\_ & 2 / 13 & 153 & 153 \\
\_Thu\_15.05.2014\_11\%3A57\%3A42\_vep\_ & 4 / 16 & 205 & 206 \\
\_Thu\_15.05.2014\_14\%3A13\%3A26\_vep\_ & 10 / 16 & 192 & 191 \\
\_Thu\_15.05.2014\_15\%3A54\%3A54\_vep\_ & 12 / 16 & 208 & 208 \\
\_Thu\_15.05.2014\_16\%3A47\%3A47\_vep\_ & 15 / 16 & 208 & 208 \\
\_Thu\_15.05.2014\_12\%3A15\%3A47\_vep\_ctr & 5 / 16 & 208 & 208 \\
\_Thu\_15.05.2014\_17\%3A22\%3A22\_vep\_ & 16 / 16 & 208 & 208 \\
\_Tue\_06.05.2014\_11\%3A25\%3A22\_vep\_ & 3 / 13 & 153 & 153 \\
\_Thu\_15.05.2014\_14\%3A13\%3A26\_vep\_ & 10 / 16 & 192 & 191
\end{tabular}
\end{center}
\caption{Number of Trials Averaged for VEP (this may slightly smaller than the number of LED on/off, since some segments might have been skipped)}
\end{table}
%% VEP 
%filename: _Tue_06.05.2014_11%3A17%3A10_vep_,    2 / 13
%Number of Trials: 153
%Number of Trials: 153
%filename: _Thu_15.05.2014_11%3A57%3A42_vep_,    4 / 16
%Number of Trials: 205
%Number of Trials: 206
%filename: _Thu_15.05.2014_14%3A13%3A26_vep_,    10 / 16
%Number of Trials: 192
%Number of Trials: 191
%filename: _Thu_15.05.2014_15%3A54%3A54_vep_,    12 / 16
%Number of Trials: 208
%Number of Trials: 208
%filename: _Thu_15.05.2014_16%3A47%3A47_vep_,    15 / 16
%Number of Trials: 208
%Number of Trials: 208
%filename: _Thu_15.05.2014_12%3A15%3A47_vep_ctr, 5 / 16
%Number of Trials: 208
%Number of Trials: 208
%filename: _Thu_15.05.2014_17%3A22%3A22_vep_,    16 / 16
%Number of Trials: 208
%Number of Trials: 208
%filename: _Tue_06.05.2014_11%3A25%3A22_vep_,    3 / 13
%Number of Trials: 153
%Number of Trials: 153
%filename: _Thu_15.05.2014_14%3A13%3A26_vep_,    10 / 16
%Number of Trials: 192
%Number of Trials: 191

\begin{table}[H]
\begin{center}
\begin{tabular}{cc|cccc}
&&\multicolumn{2}{c}{Experiment} & \multicolumn{2}{c}{Resting} \\
Filename & Index & Time Step & Length & Time Step & Length \\
\hline
\_Thu\_15.05.2014\_12\%3A08\%3A22\_ssvep\_40      & 11 / 39 & 320640  584434 & 27.4785 & 9600 301440 & 30.4000 \\
\_Thu\_15.05.2014\_12\%3A31\%3A02\_ssaep\_86      & 15 / 40 & 307603  585728 & 28.9714 & 9600 288403 & 29.0420 \\
\_Tue\_06.05.2014\_11\%3A14\%3A51\_ssvep\_40      &  5 / 39 & 320045  583839 & 27.4785 & 9600 300845 & 30.3380 \\
\_Tue\_06.05.2014\_11\%3A37\%3A22\_ssaep\_86      &  8 / 41 &  70740  585728 & 53.6446 & 9600  51540 &  4.3688 \\
\_Tue\_06.05.2014\_11\%3A23\%3A01\_ssvep\_40      &  8 / 39 & 317488  581282 & 27.4785 & 9600 298288 & 30.0717 \\
\_Tue\_06.05.2014\_11\%3A42\%3A15\_ssaep\_86      & 11 / 41 &  94848  585856 & 51.1467 & 9600  75648 &  6.8800 \\
\_Thu\_15.05.2014\_12\%3A13\%3A26\_ssvep\_ctr\_40 & 14 / 39 & 328552  592346 & 27.4785 & 9600 309352 & 31.2242 \\
\_Thu\_15.05.2014\_12\%3A26\%3A26\_ssaep\_ctr\_86 & 12 / 40 & 305364  585856 & 29.2179 & 9600 286164 & 28.8087 \\
\_Thu\_15.05.2014\_09\%3A52\%3A53\_ssvep\_40      &  2 / 39 & 328875  592669 & 27.4785 & 9600 309675 & 31.2578 \\
\_Thu\_15.05.2014\_12\%3A08\%3A22\_ssvep\_40      & 11 / 39 & 320640  584434 & 27.4785 & 9600 301440 & 30.4000 \\
\_Thu\_15.05.2014\_14\%3A20\%3A24\_ssvep\_40      & 23 / 39 & 371398  587192 & 22.4785 & 9600 304198 & 30.6873 \\
\_Thu\_15.05.2014\_16\%3A02\%3A44\_ssvep\_40      & 32 / 39 & 609145 1156050 & 56.9693 & 9600 589945 & 60.4526 \\
\_Thu\_15.05.2014\_16\%3A38\%3A47\_ssvep\_40      & 35 / 39 & 611121 1158026 & 56.9693 & 9600 591921 & 60.6584 \\
\_Thu\_15.05.2014\_12\%3A13\%3A26\_ssvep\_ctr\_40 & 14 / 39 & 328552  592346 & 27.4785 & 9600 309352 & 31.2242 \\
\_Thu\_15.05.2014\_17\%3A18\%3A01\_ssvep\_40      & 38 / 39 & 606919 1153824 & 56.9693 & 9600 587719 & 60.2207 \\
\_Thu\_15.05.2014\_10\%3A04\%3A16\_ssaep\_86      &  3 / 40 & 303843  585856 & 29.3764 & 9600 284643 & 28.6503 \\
\_Thu\_15.05.2014\_12\%3A31\%3A02\_ssaep\_86      & 15 / 40 & 307603  585728 & 28.9714 & 9600 288403 & 29.0420 \\
\_Thu\_15.05.2014\_14\%3A26\%3A54\_ssaep\_86      & 24 / 40 & 301020  585728 & 29.6571 & 9600 281820 & 28.3562 \\
\_Thu\_15.05.2014\_16\%3A12\%3A19\_ssaep\_86      & 33 / 40 & 593190 1161856 & 59.2360 & 9600 573990 & 58.7906 \\
\_Thu\_15.05.2014\_16\%3A58\%3A34\_ssaep\_86      & 36 / 40 & 595230 1161728 & 59.0102 & 9600 576030 & 59.0031 \\ \_Thu\_15.05.2014\_12\%3A26\%3A26\_ssaep\_ctr\_86 & 12 / 40 & 305364  585856 & 29.2179 & 9600 286164 & 28.8087 \\
\_Thu\_15.05.2014\_17\%3A12\%3A38\_ssaep\_86      & 40 / 40 & 595109 1161728 & 59.0228 & 9600 575909 & 58.9905
\end{tabular}
\end{center}
\caption{Start and end time step for experiment and resting state durign SSAEP and SSVEP.}
\end{table}
%[( 584434 - 320640) / 9600;
% ( 585728 - 307603) / 9600;
% ( 583839 - 320045) / 9600;
% ( 585728 -  70740) / 9600;
% ( 581282 - 317488) / 9600;
% ( 585856 -  94848) / 9600;
% ( 592346 - 328552) / 9600;
% ( 585856 - 305364) / 9600;
% ( 592669 - 328875) / 9600;
% ( 584434 - 320640) / 9600;
% ( 587192 - 371398) / 9600;
% (1156050 - 609145) / 9600;
% (1158026 - 611121) / 9600;
% ( 592346 - 328552) / 9600;
% (1153824 - 606919) / 9600;
% ( 585856 - 303843) / 9600;
% ( 585728 - 307603) / 9600;
% ( 585728 - 301020) / 9600;
% (1161856 - 593190) / 9600;
% (1161728 - 595230) / 9600;
% ( 585856 - 305364) / 9600;
% (1161728 - 595109) / 9600]

%[(301440 - 9600) / 9600;
% (288403 - 9600) / 9600;
% (300845 - 9600) / 9600;
% ( 51540 - 9600) / 9600;
% (298288 - 9600) / 9600;
% ( 75648 - 9600) / 9600;
% (309352 - 9600) / 9600;
% (286164 - 9600) / 9600;
% (309675 - 9600) / 9600;
% (301440 - 9600) / 9600;
% (304198 - 9600) / 9600;
% (589945 - 9600) / 9600;
% (591921 - 9600) / 9600;
% (309352 - 9600) / 9600;
% (587719 - 9600) / 9600;
% (284643 - 9600) / 9600;
% (288403 - 9600) / 9600;
% (281820 - 9600) / 9600;
% (573990 - 9600) / 9600;
% (576030 - 9600) / 9600;
% (286164 - 9600) / 9600;
% (575909 - 9600) / 9600]

%% SSAVEP
%filename: _Thu_15.05.2014_12%3A08%3A22_ssvep_40,    11 / 39
%Experiment: 320640 584434
%Baseline: 9600 301440
%filename: _Thu_15.05.2014_12%3A31%3A02_ssaep_86,    15 / 40
%Experiment: 307603 585728
%Baseline: 9600 288403
%filename: _Tue_06.05.2014_11%3A14%3A51_ssvep_40,    5 / 39
%Experiment: 320045 583839
%Baseline: 9600 300845
%filename: _Tue_06.05.2014_11%3A37%3A22_ssaep_86,    8 / 41
%Experiment: 70740 585728
%Baseline: 9600 51540
%filename: _Tue_06.05.2014_11%3A23%3A01_ssvep_40,    8 / 39
%Experiment: 317488 581282
%Baseline: 9600 298288
%filename: _Tue_06.05.2014_11%3A42%3A15_ssaep_86,    11 / 41
%Experiment: 94848 585856
%Baseline: 9600 75648
%filename: _Thu_15.05.2014_12%3A13%3A26_ssvep_ctr_40,    14 / 39
%Experiment: 328552 592346
%Baseline: 9600 309352
%filename: _Thu_15.05.2014_12%3A26%3A26_ssaep_ctr_86,    12 / 40
%Experiment: 305364 585856
%Baseline: 9600 286164
%filename: _Thu_15.05.2014_09%3A52%3A53_ssvep_40,    2 / 39
%Experiment: 328875 592669
%Baseline: 9600 309675
%filename: _Thu_15.05.2014_12%3A08%3A22_ssvep_40,    11 / 39
%Experiment: 320640 584434
%Baseline: 9600 301440
%filename: _Thu_15.05.2014_14%3A20%3A24_ssvep_40,    23 / 39
%Experiment: 323398 587192
%Experiment: 371398 587192
%Baseline: 9600 304198
%filename: _Thu_15.05.2014_16%3A02%3A44_ssvep_40,    32 / 39
%Experiment: 609145 1156050
%Baseline: 9600 589945
%filename: _Thu_15.05.2014_16%3A38%3A47_ssvep_40,    35 / 39
%Experiment: 611121 1158026
%Baseline: 9600 591921
%filename: _Thu_15.05.2014_12%3A13%3A26_ssvep_ctr_40,    14 / 39
%Experiment: 328552 592346
%Baseline: 9600 309352
%filename: _Thu_15.05.2014_17%3A18%3A01_ssvep_40,    38 / 39
%Experiment: 606919 1153824
%Baseline: 9600 587719
%filename: _Thu_15.05.2014_10%3A04%3A16_ssaep_86,    3 / 40
%Experiment: 303843 585856
%Baseline: 9600 284643
%filename: _Thu_15.05.2014_12%3A31%3A02_ssaep_86,    15 / 40
%Experiment: 307603 585728
%Baseline: 9600 288403
%filename: _Thu_15.05.2014_14%3A26%3A54_ssaep_86,    24 / 40
%Experiment: 301020 585728
%Baseline: 9600 281820
%filename: _Thu_15.05.2014_16%3A12%3A19_ssaep_86,    33 / 40
%Experiment: 593190 1161856
%Baseline: 9600 573990
%filename: _Thu_15.05.2014_16%3A58%3A34_ssaep_86,    36 / 40
%Experiment: 595230 1161728
%Baseline: 9600 576030
%filename: _Thu_15.05.2014_12%3A26%3A26_ssaep_ctr_86,    12 / 40
%Experiment: 305364 585856
%Baseline: 9600 286164
%filename: _Thu_15.05.2014_17%3A12%3A38_ssaep_86,    40 / 40
%Experiment: 595109 1161728
%Baseline: 9600 575909

\end{document}