### 2014-08-14 analysis/release/layout/figures.tex
Add more documented code.

\documentclass[]{article}

\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{multirow}

\pagestyle{plain}

\begin{document}

\setlength{\tabcolsep}{2pt}
\begin{figure}[H]
\begin{center}
\includegraphics[]{../figures/legend.pdf}
\hspace{0.3cm}
\begin{tabular}{c|cc|cc}
& \multicolumn{2}{c|}{Experiment} & \multicolumn{2}{c}{Control} \\
& Subject 1 & Subject 2 & Subject 1 & Subject 2 \\
\hline
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_11_57_42_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_17_10-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_12_15_47-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_25_22-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_11_57_42_off_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_17_10_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_12_15_47_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_25_22_off-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_14_26_54_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_37_22-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_12_26_26-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_42_15-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_14_20_24_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_14_51-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_12_13_26-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Tue_06_05_2014_11_23_01-crop.pdf}}
\end{tabular}
\caption{}
\end{center}
\end{figure}
\setlength{\tabcolsep}{6pt}

\clearpage

\setlength{\tabcolsep}{2pt}
\begin{figure}[H]
\begin{center}
\begin{tabular}{c|cccc|cc}
         & \multicolumn{4}{c|}{Experiment}                       & \multicolumn{2}{c}{Control} \\
         &             &             &             &             & Live         & Dead \\
\hline
Location & Basilar Tip & Mid-Basilar & VB Junction & Basilar Tip & Basilar Tip  & Basilar Tip \\
Time     & 11:57:42    & 14:13:26    & 15:54:54    & 16:47:47    & 12:15:47     & 17:22:22 \\
Time     & 0:00:00 & 2:15:44 & 3:57:12 & 4:50:05 & 0:18:05 & 5:24:40 \\ % TODO: time relative to anesthesia? relative to surgery
\hline
\raisebox{-0.5\height}{\rotatebox{90}{VEP On}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_11_57_42_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_14_13_26-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_15_54_54-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_16_47_47-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_12_15_47-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_17_22_22-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{VEP Off}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_11_57_42_off_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_14_13_26_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_15_54_54_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_16_47_47_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_12_15_47_off-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_17_22_22_off-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSAEP 86 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_12_31_02_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_14_26_54-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_16_12_19-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_16_58_34-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_12_26_26-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_17_12_38-crop.pdf}} \\
\raisebox{-0.5\height}{\rotatebox{90}{SSVEP 40 Hz}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_12_08_22_labelled-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_14_20_24-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_16_02_44-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_16_38_47-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_12_13_26-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.30]{../figures/Thu_15_05_2014_17_18_01-crop.pdf}}
\end{tabular}
\caption{}
\label{fig:position}
\end{center}
\end{figure}
\setlength{\tabcolsep}{6pt}

\clearpage

\begin{figure}[H]
\begin{center}
\begin{tabular}{cc|cc}
\raisebox{-0.5\height}{\multirow{3}{*}{
\begin{tabular}{c}
Cardiac-Aligned  Measurements \vspace{0.7cm} \\
\includegraphics[scale=0.64]{../figures/Thu_15_05_2014_14_13_26_cardiac-crop.pdf}
\end{tabular}
}} &
& Raw Trace & VEP-Aligned \\
\hline
&
\raisebox{-0.5\height}{\rotatebox{90}{Contaminated}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_14_13_26_start-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_14_13_26_cardiac_labelled-crop.pdf}} \\
&
\raisebox{-0.5\height}{\rotatebox{90}{Cleaned}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_14_13_26_end-crop.pdf}} &
\raisebox{-0.5\height}{\includegraphics[scale=0.32]{../figures/Thu_15_05_2014_14_13_26_labelled-crop.pdf}} \\
\end{tabular}
\caption{}
\end{center}
\end{figure}
\end{document}